---
title: "STAT1400 Statistics for Science"
subtitle: "Computer Lab Week 7"
output:
  html_document: 
    toc: true
  html_notebook: default
  pdf_document: 
    toc: true
    number_sections: yes
  word_document: 
    toc: true
# date: '`r format(Sys.time(), "%d %B %Y, %H:%M ")`'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(prompt=FALSE, comment=NA, message=FALSE, tidy=TRUE, error=FALSE, eval=TRUE)
```


# Learning Check

In this lab you should read through and run the code in the lab sheet and complete the lab assessment. 

When you complete this session, you will be able to:

1. work with iterations;

2. understand Boolean (True or False) values and basic logical comparisons;

3. work with conditionals.


# Copyright and UWA unit content

**Is it OK to download and share course material such as lectures, unit outlines, exam papers, articles and ebooks?**

UWA is committed to providing easy access to learning material and many of your lectures are available for online access via the Lecture Capture System (LCS), accessible through the LMS. Your unit coordinator may make their lecture recordings available to download if they wish. You are allowed to access recorded lectures in the format they are supplied on the LCS – so if they are not made available to download, you must not use any software or devices to attempt to download them.

All recorded lectures and other course material, such as presentation slides, lecture and tutorial handouts, unit outlines and exam papers, are protected under the Copyright Act and remain the property of the University. You are not allowed to share these materials outside of the LMS – for example, by uploading them to study resource file sharing websites or emailing them to friends at other universities. Distributing course material outside of the LMS is a breach of the University Policy on Academic Conduct and students found to be sharing material on these sites will be penalised. University data, emails and software are also protected by copyright and should not be accessed, copied or destroyed without the permission of the copyright owner.

Other material accessible from the LMS or via the Library, such as ebooks and journal articles, are made available to you under licensing agreements that allow you to access them for personal educational use, but not to share with others.
 
**Can I share my login details?**

No! Pheme is your key to accessing a number of UWA's online services, including LMS, studentConnect, UWA email, your Library account, and Unifi. These services hold copyright material as well as your personal information, including your unit marks, enrolment information and contact details, so it is important that you do not share the access credentials with anyone else.

https://www.student.uwa.edu.au/learning/resources/ace/respect-intellectual-property/copyright-and-uwa-unit-content


## Question 1: Understanding "Iterations"

This week, we will improve our programming skills in R. We will start with "iterations". As you can see, in R we can use the `for()` command to repeat a task multiple times; this is also referred to as a **'for loop'**. Let's start with an easy example of a for loop to get a better understanding of its structure.

```{r}
for (i in 1:10) {
    print(i)
}
```

The first line states the limits of the for loop. So 1 to 10 in the exapmle above. The second line is called "the body", which is executed once for each item of the for loop. So here we execute `print(i)`10 times for changing values of i. Note that i increases by 1 between executions of the body. 

### Exercise 1.1: For Loop Iteration

The next block of code will show you two possibilities of how a for loop can be used to calculate large sums (here: $\sum_{i=1}^{100}\, i$) or approximate series. A series is an infinite sum (for example: $\sum_{i=1}^{\infty}\,i$). Try to understand the two methods in the code cells below. Can you see how they differ? Then use this technique to calculate the sums/ approximate the series in your assessment.

```{r}
results <- c()                    # creates an empty vector
for (i in 1:100){
    results[i] <- i               # i is stored at i-th position in the vector
    }
sum(results)                      # all vector entries are being summed up
    
```

```{r}
result <- 0                      # sets initial value to 0
for (i in 1:100){
    result <- result + i        # the initial value is replaced by (itself + i)
    }
result                          # the value of "results" is printed
```



### Exercise 1.2 Weather dataset

Do you remember how we determined the warmest year (within the years 2000-2009) based on average (over months) temperature? 

```{r}
weather <- read.csv("weather.csv")
head(weather,n=24)
```

Use the template below to program a for loop which stores the average (over all 12 months) temperature for each year in a vector called 'vec'.

```{r}
number.years <- 2009-1950+1 
vec <- rep(0,number.years)
length(vec)
```

```{r}
for(i in 1:number.years){
    jan <- 12*i-11
    dec <- 12*i
    vec[i] <- mean(weather$average.temp[jan:dec])       # to be completed by you 
}

vec

mean(weather$average.temp[1:12])              # the first entry in 'vec' should be equal to this number.
```

Now answer the above questions.

## Question 2: Booleans (True or False)

Another important feature in R is called "conditional". In order to properly understand how conditionals work, we firstly need to get a greater understanding of "Booleans".  Boolean values can either be `TRUE` or `FALSE`. We get Boolean values when using comparison operators, amongst which are `<` (less than), `>` (greater than), and `==` (equal to).\
You have already seen Booleans a few times throughout past labs (remember that the `==` was helpful for separating the Maths from the Physics students), but let's take a closer look still.

Run the cells below to see examples of comparison operators in action.

```{r}
5 < 4-6
5 < 4+6
5 == 4-6
5 == 3+2
5 != 3+2
3 < 4 
(3:10)==5
```

Note the `!=` notation in the last line of the previous code block which opposes the `==` and reads as "is not equal to". We can also assign the result of a comparison operation to a variable.

```{r}
bool <- 20 -15 == 15/3
bool
```
```{r}
AGE <- c(42,36,25,20,55,19,25,35,19,24)
summary(AGE)
```




Vectors are also compatible with comparison operators. The output is a vector of Boolean values.

```{r}
c(1, 5, 7, 8, 3, -1) > 3
```

Also remember from the lecture, that in R the `&&` symbol represents the "AND condition" and that both sides of `&&`  have to be true for an AND-condition to be true.

```{r}
5 < 7 && 4 > 3
5 < 7 && 4 > 5
```

From the code below, can you guess what condition is represented by the `||` symbol? OR

```{r}
5 < 7 || 4 > 3
5 < 7 || 4 > 5
5 < 3 || 2 > 3
```

If we compare vectors of Booleans, `&&` and `||` will only compare the first element of both vectors. Instead we have to use `&` and `|` if we want to compare the vectors element by element. For example:

```{r}
x <- c(1,2,3,4)
y <- c(0,2,4,-1)

x<y & x>y

x<y | x>y
```


## Question 3: Conditionals

Let us now get back to the earlier mentioned "conditional". A conditional statement is made up of multiple lines that allow R to choose from different alternatives based on whether some condition is 'TRUE' or 'FALSE'. It always begins with an "if header", which is a single line consisting of "if" and a condition in parentheses. This line is then followed by the "body".
The body is only executed if the condition of the "if-header" evaluates to 'TRUE'. If the condition evaluates to 'FALSE', then the body of that "if header" is skipped. See the following code for clarification.

```{r}
if (3 < 5){
    print ("3 is smaller than 5")
} else{
    print ("3 is not smaller than 5")
}


if (7 < 4){
    print ("7 is smaller than 4")
} else{
    print ("7 is not smaller than 4")
}
```
```{r}
if (3<4){
  print("3 is smaller than 4")
} else{
   print("four is smaller than 3")
}   
```
### 3.1 Functions

Conditional statements often appear within the body of a function. 

We have already seen many examples of functions, like `summary()`, `print()` and `mean()` to name a few. These are in-built functions which we can readily access. However, we can also build our own functions. In general, a function performs a task based on its arguments. The `mean()` function for instance calculates and returns the average of its argument, which has to be a vector of numerical values.

If we would like to create our own function we need to use this basic syntax:

```
function_name <- function(arg_1, arg_2, ...) {
   function body 
}
```

As you can see a function is created by assigning the keyword 'function' followed by a set of arguments to the variable 'function_name'. If we want to call the function later, we simply type 

```
function_name(arg_1, arg_2,...)
```

The code `mean(c(2,5,4))` is an example of that where 'mean' is the function name and 'c(2,5,4)' is its first (and only) argument. 

The following lines of code present a function which accepts a vector as input and calculates the sum of the squared elements of that vector. 

```{r}
sum.of.squares <- function(x){
    return(sum(x^2))
}
   
sum.of.squares(c(1,2,3))
```

Getting back to conditionals which often appear within the body of a function to express alternative behavior depending on the argument of the function: Study the following example for clarification. Also notice the `else if` command which adds another condition to our code. The "body" of that "else if" condition is only executed if the original condition (here x < y) is 'FALSE'. The final `else` is only reached if all previous conditions evaluate to 'FALSE'.

```{r}
fn.1 = function(x,y) {
    if( x < y) {
        cat(x,"is smaller than", y, "\n")
    } else if (x > y) {
        cat(y,"is smaller than", x, "\n")   
    } else {
        cat(x,"is equal to", y, "\n")
    }
}

fn.1(3,5)
fn.1(4,2)
fn.1(6,6)
```

The general format of a multi-clause conditional statement looks like the one above.

   - There is always exactly one "if header", but there can be any number of "else if" clauses. R will evaluate the
"if condition" first. Then, in case the previous expression was 'FALSE' the next "else if condition" is evaluated and so on until one is found that is 'TRUE'. Then R executes the corresponding "body".

   - The "else clause" is optional and often serves as a collection of all remaining cases which weren't part of any "if" or "else if condition" yet. Hence, when an "else" clause is provided, its "body" is executed
only if none of the header expressions of the previous clauses are 'TRUE'. The "else" clause is always the last clause or doesn't appear at all. A code template looks like this:

```
if (condition 1){
    if-body 
} else if (condition 2) {
    else-if-body 
} else if (condition 3) {
    else-if-body 
}  ...
  else {
    else-body
}
```
```{r}
round(pexp(10,rate = 0.2),4)
qexp(0.9)
```
```{r}
result <- 0
for (i in 1:99) {
  result <- result + i^2 - 2*i
}
result
```
```{r}
result <- 0
for (i in 11:25) {
  result <- result + (i^3)*log(4)
}
round(result,3)
```
```{r}
x <-c(-3,0,1,-3)

y <- c(-1,-1,1,-2)

x==y & x<y
```
```{r}
x <-c(-3,0,1,-3)

y <- c(-1,-1,1,-2)

x<y | x>y
```




### Exercise 3.1: Monkey bet

Imagine betting on a monkey throwing a dart (a dartboard consist of 20 equally sized panels labeled 1 to 20. For the purpose of this exercise we ignore the fact that the centre of the board consists of two additional panels). The rules of the game are:

1) if the monkey hits the panel '1', you lose 1 dollar\
2) if the monkey hits an even number nothing happens\
3) if the monkey hits an odd number between 3 and 9 (including both), you win 2 dollars\
4) if the monkey hits an odd number between 11 and 17 (including both), you win 4 dollars\
5) if the monkey hits 19, you win 10 dollars


Your task now is to write a function that returns your winnings. 

Hint: In R we can test if an integer $a$ is divisible by $b$ via `a%%b`, which will state the remainder. See examples below.

```{r}
24 %% 2
47 %% 5
```
```{r}
monkey.bet = function(x) {
  if (x>20 | x<1) {
  stop ("number is not a dart panel")
} else if (x == 1) {
    return(-1)
} else if (x %% 2 == 0) {
    return(0) 
} else if (x == 19) {
    return(10)
} else if (3 < x && x <9 && x %% 2 != 0){
    return(2)
} else {
    return(4)
}
}
 monkey.bet(5) 
```
```{r}
set.seed(1400)
pay <- c()
for (i in 1:20) {
  x <- sample(1:20,size = 1,replace = T)
  pay[i] <- monkey.bet(x)
}
mean(pay)
var(pay)
```








```{r}
set.seed(1400)
results <- c()
for (i in 1:10000) {
  x <- sample(1:20,size = 1)  
  results[i] <- monkey.bet(x)
}
var(results)
mean(results)
tab <- table(results)
tab
prop.table(tab)
round(0.0967+ 0.2961+ 0.0516,3)
round(1-0.0518,3)
p = c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18,19,20)
n = 20
px = 1/20
ex = sum(p*px)
ex
variance = sum(p - ex)^2*px
variance
```
```{r echo=TRUE}
set.seed(1400)
A<-0  ## step 0
for(i in 1:10000){
u1<-runif(1); u2<-runif(1)  ## step 1
if(u1<.9 || u2<.7){   ##  || is OR  && is AND  step 2
A<-A+1
}
}
A
Phat=A/10000
Phat
```


Now insert the missing code pieces in the code block below to create a function called "monkey.bet", which returns your winnings for one throw of a dart. 

```{r eval=FALSE, include=FALSE}
# THIS CODE BLOCK WILL HAVE TO BE MANIPULATED BY YOU TO WORK!
# To do so, replace all sections which include "...j..." with the missing code for j=1,2,3,4.

# Complete all if-, else if- and else clauses.


monkey.bet = function(x) {
# Returns payout if the dart lands on panel x.
    if (...1...) {
        return (-1)
    }
    else if (...2...) {
        return (0)
    }
    else if (...3...) {
        return (2)
    }
    else if (...4...) {
        return (4)  
    }
    else if (x>20 | x<1) {stop ("number is not a dart panel")}
    else  {
        return (10)   
    }
}
```

Use the code below to check your "monkey.bet" function.

```{r eval=FALSE, include=FALSE}
results <- c()
for (i in 1:20) {
    results[i] <- monkey.bet(i)
}
results
tab <- table(results)
tab
prop.table(tab)
var(results)
```

```{r}
set.seed(1400)
monopoly1 <- sample(1:6, size=100, replace=T)+sample(1:6, size=100, replace=T)

set.seed(1400)
d1 <- sample(1:6, size=100, replace=T)
d2 <- sample(1:6, size=100, replace=T)
monopoly2 <- d1+d2

X    <- 2:12
prob <- (c(1,2,3,4,5,6,5,4,3,2,1)/36)
set.seed(1400)
monopoly3 <- sample(X, size=100, replace=T, prob=prob)
mean(monopoly3)
TAB <- table(monopoly3)
TAB
prop <- prop.table(TAB)
prop
round(0.03+0.05+0.1,3)
```
```{r echo=TRUE}
set.seed(1400)
A<-0  ## step 0
for(i in 1:10000){
u1<-runif(1); u2<-runif(1)  ## step 1
if(u1<.3 || u2<.5){   ##  || is OR  && is AND  step 2
A<-A+1
}
}
A
Phat=A/10000
Phat
```
```{r echo=TRUE}
set.seed(1400)
A<-0  ## step 0
for(i in 1:10000){
u1<-runif(1); u2<-runif(1)  ## step 1
if(u1<.9 || u2<.7){   ##  || is OR  && is AND  step 2
A<-A+1
}
}
A
Phat=A/10000
Phat
```
```{r}
pexp(10,rate = 0.2)-pexp(5,rate = 0.2)
```
```{r}
pexp(4,rate = 1.5)-pexp(2,rate = 1.5)
```
```{r}
# complete this code to plot the two histograms.
# Make sure you remove eval=FALSE, include=FALSE

set.seed(1400)

n = 50
original = rexp(n,1.5)

n=50
m=500
set.seed(1400)
sample_means = c()
for (i in 1:m) {
  sample_means[i] = sum(rexp(n,1.5))/n
}
mean(sample_means)
var(sample_means)
mean(original)
var(original)
summary(original)
IQR(original)
par(mfrow=c(1,2))
hist(original,freq=TRUE,col=2, main="original sample")
hist(sample_means,freq=TRUE,col=2, main="sample mean")
```
```{r}
x <- c(1,2,3,4)
y <- c(0,2,4,-1)

x<y & x>y

x<y | x>y
```

```{r}
x <-c(-3,0,1,-3)
y <- c(-1,-1,1,-2)
x==y & x<y
```
```{r}
(0.8172454)^2
```









* What happens when you call your function with an argument of "21" instead?

* Why does this happen?

To assure that the function only accepts arguments which are elements of the sample space add the command line below into your monkey.bet function as the first "if-header". 

```{r eval=FALSE, include=FALSE}
if (x>20 | x<1) {stop ("number is not a dart panel")}
```

If you now call the function with "21" again, you should receive an error message.

### Exercise 3.2 Sum of Squares

We can use our knowledge about conditionals in order to find how many summands are needed for a series to get above a certain threshold. For example let us take the series $\sum_{i=1}^{n} i^2$ and the number of summands needed to exceed the threshold of 1000. Since the sum starts at $i = 1$ finding the number of summands is equal to finding the value of $n$, for which this series is greater than $1000$ for the first time. (Note that this changes once the sum does not start at $i = 1$.)
To find out, study the following lines of code. Notice that the `break` command exits the for loop.

```{r}
result <- 0                     # sets initial value to 0
counter <- 0                    # sets the summand counter to 0

for (i in 1:100){
    result <- result + i^2      # the initial value is replaced by (itself + i^2)
    counter <- counter + 1      # the summand counter increases by 1.
    if (result > 1000) {        # conditional
        print(counter)          # prints the number of summands that were needed to get past the threshold
        break                   # exists the for loop.
    }
 }
```

As you can see by the output, 14 summands are needed (i.e. $1^2 + 2^2 + 3^2 +...+ 14^2$) until the sum is greater than 1000. Now use the knowledge you have gathered throughout this lab sheet to manipulate the code above such that it will help you answer similar questions. 

```{r}
result <- 0 # starts at zero(initial value)
counter <- 0 # starts at zero (initial value)
for (i in x : y) {
  result <- result + calculation formula
  counter <- counter +1
  if (result condition){
    print(counter)
    break 
  }
}
```

### Exercise 3.3: United Airlines data

We will now bring all of the concepts together in one final exercise. First import the datafile below. It contains information on a total of 13825 United Airlines domestic flights departing from San Francisco in the summer of 2015.

```{r}
united <- read.csv("summer.csv")
head(united)
nrow(united)
```

It is your job now, to add a filter variable called 'Delayed' to the dataframe, which will indicate whether a flight was on time (if delay is 0 or negative) or delayed (if delay is positive). First, create a vector with the name 'Delayed', that will contain this information and later be added to the dataframe.

```{r}
Delayed <- rep(0,13825)
length(Delayed)
for (i in 1:length(Delayed)) {
  if (united$Delay[i] > 0) {
    Delayed[i] = 1
  } else {
    Delayed[i] = 0
  }  
}
```

In the cell block below, write a small R script yourself as follows: Use a for loop to go through all 13285 flights. For each flight, use a conditional to check whether the flight was delayed or not. If it was delayed, store a TRUE or a 1 at that position of the 'Delayed' vector. If it was not delayed, store a FALSE or a 0.

Note: Faster code routines exist of course. For the purpose of this exercise it is not crucial to stick to the above outlined routine.



Run the next block of code to add the vector to the dataframe. Then, determine the proportion of delayed flights.
```{r}
result <- 0                     # sets initial value to 0
counter <- 0                    # sets the summand counter to 0

for (i in 1:99){
    result <- result + i^2 - i      # the initial value is replaced by (itself + i^2)
    counter <- counter + 1      # the summand counter increases by 1.                      # exists the for loop
}
result
```
```{r}
result <- 0
for (i in 11:25) {
  result <- result + (i^2)*log(4)
}
result
```
```{r}
set.seed(1400)
monopoly <- sample(1:6,size = 100,replace = T)+sample(1:6,size = 100,replace = T)
set.seed(1400)
x <- 2:12
prob <- (c(1,2,3,4,5,6,5,4,3,2,1)/36)
set.seed(1400)
sampling <- sample(x,size = 100, replace = T,prob = prob )
mean(sampling)
summary(sampling)
```
```{r}
result <- 0
for (i in 11:25) {
  result <- result + (i^2)*log(4)
}
result
```





```{r}
united.extended <- cbind(united,Delayed)
head(united.extended)
```
```{r}
tab <- table(united.extended$Delayed)
tab
prop.table(tab)
```


## Question 4 Simulation using random uniform number runif()

4.1 Suppose we have two independent devices which function with probabilities 0.6 and 0.7, respectively. 

            - What is the probability that at least one device functions? 
  
Calculate the exact (theoretical) probability and the approximate probability through simulation of 10000 iterations (use set.seed(1400)).
  
4.2 *Flipping a biased coin.* Consider the following game: a coin is flipped 25 times, winning $1 each time it lands heads (H) and losing $1 each time it lands tails (T). 

           - Unfortunately, the coin is biased in such a way that $P(H)=0.4$ and $P(T)=0.6.$ 

           - What is the probability of Head, which means more money at the end of the game than at the beginning? 

Use simulation using uniform RNG to find out P(H).

*Solution*

- Each run of the simulation requires 25 “random” objects: the results of the 25 coin tosses.

- We need to keep track of how much money we’ve won or lost at the end of the 25 tosses.

- Let $A={Head}$, and use the following pseudo-code:

0. Set a counter for the number of times A occurs to zero.

Repeat $n$ times:

1. Set your initial dollar amount to zero.

2. Generate 25 random numbers $u_1, . . ., u_{25}.$

3. For each $u_i < .4,$ heads was tossed, so add 1 to your dollar amount. For each $u_i \ge 0.4,$ the flip was
tails and 1 is deducted.

4. If the final dollar amount is positive (i.e., $1 or greater), add 1 to the count of occurrences for A.

Once the $n$ runs are complete, then 

$$\hat{P}(A)= \frac{number \ of \  times \ A \ occurs}{number \ of \ runs}=\frac{n(A)}{n}.$$

The R code for simulation is provided below, you need to correctly code #1, #2, #3, #4 based on the information above:

```{r eval=FALSE, include=FALSE}
set.seed(1400)
A <- 0
n <- 25
for (i in 1:n){
dollar<-0
for (j in 1:25){
u<-runif(1)
if (u< #3){
dollar<-dollar+1
}
else{dollar<- #4}
}
if (dollar>0){
A<-A+1
}
}
phat=A/n
phat
```


```{r}
0.6*0.7 + 0.6*0.3 + 0.4*0.7
```

## Question 5 Ricker model (Taken from Douglas ET AL Ch 7, see Week 1 Introduction to R bookdown)

<https://intro2r.com/> The e-book

You are a population ecologist for the day and wish to investigate the properties of the Ricker model.

The Ricker model is defined as:

$$N_{t+1}=N_t \exp[r(1−N_t K)]$$

 

where $N_t$ is the population size at time $t, r$ is the population growth rate and $K$ is the carrying
capacity. 

Write a function to simulate this model so you can conveniently determine the effect of changing $r$ and the initial population size $N_0.$ 

K is often set to 100 by default, but you want the option of being able to change this with your function. So, you will need a function with the following arguments:

      - nzero which sets the initial population size, 
      - r which will determine the population growth rate, 
      - time which sets how long the simulation will run for and 
      - K which we will initially set to 100 by default.
