---
title: "STAT1400 Statistics for Science"
subtitle: "Computer Lab Week 8"
output:
  html_document: 
    toc: true
  html_notebook: default
  pdf_document: 
    toc: true
    number_sections: yes
  word_document: 
    toc: true
# date: '`r format(Sys.time(), "%d %B %Y, %H:%M ")`'
---

```{r}
knitr::opts_chunk$set(prompt=FALSE, comment=NA, message=FALSE, tidy=TRUE, error=FALSE, eval=TRUE)
```


# Learning Check

In this lab you should read through and run the code in the lab sheet and complete the lab assessment. 

When you complete this session, you will be able to:

1. understand the important steps in an hypothesis test;
2. understand the central limit theorem and its importance in statistical inference;
3. test hypotheses about a single population proportion.
4. understand error types in hypothesis testing;
5. construct a confidence interval for a single population proportion;


## Question 1: Central Limit Theorem

The first column of the dataset *central.csv* consists of random numbers from the exponential distribution with a mean of 1 and standard deviation of 1 as the original population.

Every number in the second column (labelled $n=5$) of the dataset *central.csv* file is the mean of 5 numbers that have each been sampled from this same population. 

Every number in the third column (labelled $n=10$) is the mean of 10 numbers that have each been sampled from this same population, and so on for the $n=50$ and $n=100$ columns as well. 
```{r}
central <- read.csv("central.csv")
central
```

(a) Explore the distributions for each variable (n=1,5,10, 50,100) by providing the following: 

            - Sample Means 
            - Sample Std Dev
            - Sigma/sqrt(n)
            - Histogram and QQ plots
```{r}
mean(central$N.1)
sd(central$N.1)
sqrt(central$N.1)
hist(central$N.1)
```       
```{r}
colMeans(central)
col.SD <- apply(central, 2, sd)
col.SD
```








```{r}
mean(central$N.5)
sd(central$N.5)
sqrt(central$N.5)
hist(central$N.5)
```     
```{r}
mean(central$N.50)
sd(central$N.50)
sqrt(central$N.50)
hist(central$N.50)
```     
```{r}
mean(central$N.100)
sd(central$N.100)
sqrt(central$N.100)
hist(central$N.100)
```     
(b) Describe what happens to the shape, location and spread of the sampling distribution of the sample mean compared to the original or parent distribution as the sample size increases.
   shape skewed to left to symmertic
   spread decrease
   location
(c) Use the histograms and normal quantile plots for each variable to assess the normality of each variable.
```{r}
qqnorm(central$N.1)
qqline(central$N.1)
qqnorm(central$N.5)
qqline(central$N.5)
qqnorm(central$N.50)
qqline(central$N.50)
qqnorm(central$N.100)
qqline(central$N.100)
```
(d) What would happen if the original or parent distribution was different? Eg what if it were the uniform distribution or the normal distribution?
no change
(e) This exercise illustrates the central limit theorem. State the central limit theorem.
the distribution of sample means approximately a normal distribution as the sample size gets larger, regardless of the population's distribution.
```{r}
colMeans(central)
```
```{r}
col.sd <- apply(central, 2, sd)
col.sd
```
```{r}
col.sample.sd <- 1/sqrt(c(1,5,10,50,100))
col.sample.sd
```
```{r}
hist(central$N.1)
abline(v=1, col= 3)
qqnorm(central$N.1)
qqline(central$N.1)
hist(central$N.100)
abline(v=1, col= 3)
qqnorm(central$N.100)
qqline(central$N.100)
```


## Question 2 Cellophane flowers of yellow and green

A plant geneticist has raised 100 progeny from a cross that is hypothesized to result in a 3:1 phenotypic ratio of yellow flowered to green flowered plants. A ratio of 84 yellow:16 green is observed.

(a) Write down the null and alternative hypotheses.
let p equals to the proportion of yellow flowered
null :the phenotypic ratio of yellow flowered is 3/4
alternative :the phenotypic ratio is not equal to 3/4
(b) How many flowers of each type would you expect if the null hypothesis is true?
yellow 75 green 25
(c) Calculate the appropriate test statistic and corresponding p-value.
p^ = 84/100 p0 = 3/4 n = 100
```{r}
p0 = 3/4
phat = 84/100
n = 100
se.p=sqrt(p0*(1-p0)/n)
se.p
test.stat=(phat-p0)/(se.p)
test.stat
```
(d) What conclusion can be made?
```{r}
1-pnorm(2.078461) + pnorm(-2.078461)
pnorm(-2.08)*2
```
the p-value is very small so we reject the null
## Question 3: Confidence Intervals - single population proportion

What is the proportion of people in the Australian population who
are left handed? A sample of 90 people is surveyed. 16 state that
they are left handed.

(a) Using the formula from the notes, calculate a 95\% confidence interval for the proportion of Australians who are left handed.

(b) Create an R function of sample size $n$, the number of success $x$ and a confidence level $\alpha$ to calculate $(1-\alpha)\%$ confidence interval. Check if the function has correctly calculate the results in (a).

```{r}
qnorm(0.975)
qnorm(0.025)
```
```{r}
## fill in below
prop.ci.fn <- function(n=90,x=16,alpha=0.05){  ### n sample size, x number of success, CI 1-alpha
  phat=x/n
  se.p=sqrt(phat*(1-phat)/n)
  znorm=qnorm(1-alpha/2)
  me=znorm * se.p
  lower=phat - me
  upper=phat + me
  return(c(phat, se.p, znorm, me, lower, upper))
}

prop.ci.fn(90, 16, 0.05)
prop.ci.fn(100, 30, 0.05)
```


(c) What assumptions have you made in calculating this confidence interval?


## Question 4: Adults and video games

A PriceWaterhouseCooper report estimates that the U.S. video game market was approximately \$8.6 billion in 2007 and is expected to increase at an annual rate of 6.3\% through 2012. Who actually plays
video games? A survey, conducted by Princeton Survey Research International, reports that over half of American adults aged 18 and over play video games. The survey used a nationally representative
sample of 2054 adults. Of the total, 1063 adults said they played video games. 

(a) Estimate the proportion of adults in the population who play video games.

(b) Calculate the 90\% confidence interval for the proportion of adults in the population who play video games.

(c) Suppose that there is a new claim that the proportion of adults in the population who play video games is more than 50\%. Conduct the appropriate steps for hypothesis testing to make a conclusion. Use a significance level of 10\%.

### Exercise

Can you write an R function for the hypothesis testing in a similar way as the CI?
```{r}
inter = function(n,x,alpha){
phat = x/n
znorm = qnorm(1-alpha/2)
ph = sqrt(x/n*(1-x/n)/n)
me = znorm * ph
upper = phat + me
lower = phat - me
return(c(phat, se.p, znorm, me, lower, upper))
}
inter(50,6,0.01)
inter(50,38,0.1)
inter(200,24,0.1)
inter(200,24,0.01)
```
```{r}
qnorm(0.95)

```

```{r}
oneside = function(n, y, p){
  phat = y/n
  z = (phat - p)/sqrt(p*(1-p)/n)
  znorm = 1- pnorm(z)
  return(c(phat, z, znorm))
}
oneside(2054, 1063, 0.5)
oneside(50, 38, 0.7)
oneside(50,6,0.1)
oneside(724,72,0.07)
oneside(200,24,0.1)
## Question 5 Error Types: Effectiveness of new drug
```
 A pharmaceutical company is testing a new drug for acne. Side-effects for the new drug may include vomiting, nausea and/or fever. Subjects are randomly assigned to a control group (placebo) and treatment group (new drug). Proportion of subjects in each group whose acne is reduced  is calculated and compared between groups.
```{r}
pnorm(0.9258)
```
(a) State the null and alternative hypotheses.
null: P1 = p2 p1 is not equal to p2 p1<p2 0r p1>p2
(b) What is a Type I error and a Type II error in this context?
one it actually does damange to patients but they think it is not damaging
two it is effective but they think not
(c) Explain the consequence of committing a Type I error and Type II error.

(d) Which error type appears more serious and why?
type one 
(e) How strong should the evidence be before rejecting the null hypothesis in this example? In other words, assuming that this test is to be carried out at a fixed significance level, what would be an appropriate level?

```{r}
# Load the ggplot2 library to access the diamonds dataset
library(ggplot2)

# Take an initial random sample of size 1000 from the diamonds dataset
set.seed(123)  # Setting seed for reproducibility
initial_sample <- diamonds[sample(nrow(diamonds), 1000, replace = FALSE), ]

# Initialize a vector to store the medians from each bootstrap sample
bootstrap_medians <- numeric(1000)

# Perform bootstrap sampling 1000 times
for (i in 1:5000) {
  # Sample with replacement from the initial sample
  bootstrap_sample <- initial_sample[sample(nrow(initial_sample), replace = TRUE), ]
  
  # Calculate the median price of the bootstrap sample
  bootstrap_medians[i] <- median(bootstrap_sample$price)
}

# Calculate the 90% confidence interval of the medians
quantile(bootstrap_medians, probs = c(0.005, 0.995))
```


















